The Astropy Project

### 2017-11-20 main.tex
Group ascii and fits items

We summarize key new features or updates to \package{astropy.io} below,
organized by sub-subpackage.

\inlinecomment{Tom R}{is a sub-subpackage a thing, or is it just 'subpackage' all the way down?}

\subsubsection{Enhanced CSV format}

One of the problems when storing a table in an ASCII format is
preserving table meta-data such as comments, keywords and column data
types, units, and descriptions. Using the newly defined \emph{Enhanced
Character Separated Values} (ECSV,  \citet{ape6}) format it is now
possible to write a table to an ASCII-format file and read it back
with no loss of information. The ECSV format has been designed to be
both human-readable and compatible with most simple CSV readers.

\subsubsection{\texttt{.ascii}: fast readers and writers}

The \package{astropy.io.ascii} module now includes a significantly
faster Cython/C engine for reading and writing ASCII files. This is
available for the most common formats.

%following formats: \texttt{basic},
%      \texttt{commented\_header}, \texttt{csv}, \texttt{no\_header},
%      \texttt{rdb}, and \texttt{tab}.
On average the new engine is about
4 to 5 times faster than the corresponding pure-\python
implementation, and is often comparable to the speed of the
\package{Pandas} \citep{pandas} ASCII file interface.  The fast
reader has parallel processing option that allows harnessing
multiple cores for input parsing to achieve even greater speed
gains.

By default, \texttt{read()} and \texttt{write()} will attempt to use
the fast C engine when dealing with compatible formats. Certain
features of the full read / write interface are not available in the
fast version, in which case the pure-\python version will automatically
be used.

\subsubsection{\texttt{.ascii}: HTML tables}

The \package{astropy.io.ascii} subpackage now provides the capability
to read a table within an HTML file or web URL into an astropy
\texttt{Table} object. This requires the \package{BeautifulSoup4}
package to be installed.  Conversely a \texttt{Table} object can now
be written out as an HTML table.

\subsubsection{\texttt{.fits}: command-line scripts (\texttt{fitsheader}, \texttt{fitsinfo})}

The \package{astropy.io.fits} subpackage now provides command-line
scripts for printing a summary of the HDUs in a FITS file(s)
(\texttt{fitsinfo file.fits}) and for printing the header information
to the screen in a human-readable format (\texttt{fitsheader
file.fits}).

\subsubsection{\texttt{.fits}: lazy loading }

@@

\inlinecomment{Tom R}{too detailed? I don't think we need to mention every single performance enhancement, right?}

\subsubsection{\texttt{.misc}: YAML serialization}

The new \package{astropy.io.misc.yaml} module allows converting
astropy objects into a standard YAML format.
This can be used beyond \texttt{astropy.io.misc.yaml}: for example, to
serialize the metadata of tables before saving to other formats like
HDF5.

>>>>

% We summarize key new features or updates to \package{astropy.io} below,
% organized by sub-subpackage.

% \inlinecomment{Tom R}{is a sub-subpackage a thing, or is it just 'subpackage' all the way down?}

\subsubsection{ASCII}
% \subsubsection{Enhanced CSV format}

One of the problems when storing a table in an ASCII format is
preserving table meta-data such as comments, keywords and column data
types, units, and descriptions. The newly defined \emph{Enhanced
Character Separated Values} (ECSV,  \citet{ape6}) format makes it
possible to write a table to an ASCII-format file and read it back
with no loss of information. The ECSV format has been designed to be
both human-readable and compatible with most simple CSV readers.

% \subsubsection{\texttt{.ascii}: fast readers and writers}

The \package{astropy.io.ascii} subpackage now includes a significantly faster
Cython/C engine for reading and writing ASCII files. This is available for the
most common formats.  On average the new engine is about 4 to 5 times faster
than the corresponding pure-\python implementation, and is often comparable to
the speed of the \package{Pandas} \citep{pandas} ASCII file interface.  The
fast reader has parallel processing option that allows harnessing multiple
cores for input parsing to achieve even greater speed gains.  By default,
\texttt{read()} and \texttt{write()} will attempt to use the fast C engine
when dealing with compatible formats. Certain features of the full read
/ write interface are not available in the fast version, in which case the
pure-\python version will automatically be used.

% \subsubsection{\texttt{.ascii}: HTML tables}

The \package{astropy.io.ascii} subpackage now provides the capability
to read a table within an HTML file or web URL into an astropy
\texttt{Table} object. Conversely a \texttt{Table} object can now
be written out as an HTML table.

\subsubsection{FITS}
% \subsubsection{\texttt{.fits}: command-line scripts (\texttt{fitsheader}, \texttt{fitsinfo})}

The \package{astropy.io.fits} subpackage now provides command-line
scripts for printing a summary of the HDUs in a FITS file(s)
(\texttt{fitsinfo}) and for printing the header information
to the screen in a human-readable format (\texttt{fitsheader}).

% \subsubsection{\texttt{.fits}: lazy loading }

@@

% \inlinecomment{Tom R}{too detailed? I don't think we need to mention every single performance enhancement, right?}

% \subsubsection{\texttt{.misc}: YAML serialization}

% The new \package{astropy.io.misc.yaml} module allows converting astropy
% objects into a standard YAML format.  This can be used beyond
% \texttt{astropy.io.misc.yaml}: for example, to serialize the metadata of
% tables before saving to other formats like HDF5.
